### 2012-12-07 inst/examples/gaussian-process-basics.Rmd
cleaner plots, remove kernlab comparison

```{r include=FALSE}
opts_knit$set(upload.fun = socialR::notebook.url)
```

@@

```{r}
fig2a <- ggplot(dat,aes(x=x,y=value)) +
  geom_rect(xmin=-Inf, xmax=Inf, ymin=-2, ymax=2, fill="grey80") +
  geom_line(aes(group=variable)) +   theme_bw() +
  scale_y_continuous(lim=c(-3,3), name="output, f(x)") +
  xlab("input, x")
fig2a
```

@@

fig2c <- ggplot(dat,aes(x=x,y=value)) +
  geom_ribbon(data=Ey, aes(x=x,y=y, ymin=ymin, ymax=ymax), fill="grey80") + # Var
  geom_line(aes(color=variable)) + #REPLICATES
  geom_line(data=Ey,aes(x=x,y=y), size=1) + #MEAN
  geom_point(data=obs,aes(x=x,y=y)) +  #OBSERVED DATA
  scale_y_continuous(lim=c(-3,3), name="output, f(x)") +
  xlab("input, x") 
fig2c + opts(panel.background = theme_rect(fill = "transparent"), 
        plot.background = theme_rect(fill = "transparent"))
```

@@

## A comparison to built-in R utilities

Some basic Gaussian process tools are provided in the `kernlab` package,

```{r}
library(kernlab)
```

through the `gausspr` function,

```{r}
gp <- gausspr(obs$x, obs$y, kernel="rbfdot", kpar=list(sigma=1/(2*l^2)), fit=FALSE, scaled=FALSE, var=0.8)
```

Where `rbfdot` indicates the Gaussian (Radial basis function) kernel we use above, with parameter `sigma=1/(2l^2)` (note that gausspr defines the kernel parameter differently, see:

```{r}
gp@kernelf@.Data
```
note this takes vectors `x`, `y` and returns a scalar, not the covaraiance matrix).  The function will try to estimate hyperparameters such as `sigma` if they are not given unless `fit=FALSE`.  The data is also scaled to 0 mean, unit variance unless `scaled=FALSE` is set.

We can then use the estimated process to infer values on a specified grid through `predict`, e.g. for the example above,

```{r}
y_p <- predict(gp, x_predict)
dat2 <- data.frame(x = x_predict, variable = "Y_p", value = y_p)
```

```{r}
ggplot(dat2, aes(x, value)) + geom_point() + geom_line(data=Ey,aes(x=x,y=y), size=1) 
```

### Fitting hyperparameters by maximum likelihood

If we do not specify `fit=false` then `gausspr` will estimate the hyperparamters by maximum likelihood:

```{r}
gp_fit <- gausspr(obs$x, obs$y, kernel="rbfdot", kpar=list(sigma=1/(2*l^2)), scaled=FALSE, var=0.8)
```

>>>>

```{r include=FALSE}
opts_knit$set(upload.fun = socialR::notebook.url)
theme_update(plot.background = element_rect(fill = "transparent",colour = NA))
```

@@

```{r}
fig2a <- ggplot(dat,aes(x=x,y=value)) +
  geom_rect(xmin=-Inf, xmax=Inf, ymin=-2, ymax=2, fill="grey80") +
  geom_line(aes(group=variable)) +  
  scale_y_continuous(lim=c(-3,3), name="output, f(x)") +
  xlab("input, x")
fig2a
```

@@

fig2c <- ggplot(dat,aes(x=x,y=value)) +
  geom_ribbon(data=Ey, aes(x=x,y=y, ymin=ymin, ymax=ymax), fill="grey80") + # Var
  geom_line(aes(color=variable)) + #REPLICATES
  geom_line(data=Ey,aes(x=x,y=y), size=1) + #MEAN
  geom_point(data=obs,aes(x=x,y=y)) +  #OBSERVED DATA
  scale_y_continuous(lim=c(-3,3), name="output, f(x)") +
  xlab("input, x") 
fig2c
```

@@
